---
title: "RPart1_Homework"
author: "NYC Data Science Academy"
output: html_document
---
## Question #1	

Compound interest can be computed using the formula
$$A = P (1+\frac{r}{100}) ^n$$
where P is the original money lent, A is what it amounts to in n years at R percent per year interest. Write an R command to calculate a vector of numbers indicating the amount of money owed (A) after n years, where n ranges from 1 to 15 in yearly increments. The original amount lent is 5000 dollars (P) and the interest rate remains constant throughout the period at 11.5% (r). 

```{r Question 1}
n = 1:15
P = 5000
r = .115

interest = function(n){
  5000*(1+(11.5/100))^n
}
a = sapply(n, interest)
print(a)

```

## Question #2

Assume that we have collected the heights and weights of four people. The heights (in cm) are 180, 165, 160, 193; the respective weights (in kg) are 87, 58, 65, 100. Create two vectors, `height` and `weight`, using the data. Body mass index (BMI) is defined as
$$BMI = \frac{mass_{kg}}{height_{m}^2}$$
Write an R command to make a vector calculating the BMI values for the four people. Be careful of the units! As a challenge, use the height vector to make a boolean vector named `tall` of the heights above 6 feet.

```{r Question 2}
m = c(87, 58, 65, 100)
h = c(180, 165, 160, 193)
bmi = function(m,h){
  m/(h/10)
}
mapply(bmi, m, h)




```


## Question #3:

1. From your RStudio, import the built-in cars dataset by running `data(cars)`.

2. Print the first 5 lines from cars.

3. Randomly generate a vector as long as the the number of rows in cars containing elements NY, CA or CT. Call the vector `state`. Run the code `set.seed(0)` on the line above your vector. This makes your results reproducible (anybody who runs the code `set.seed(0)` on their randomized vector will end up with the same random vector you generated). 

4. Add state to the data frame cars as a new column. Again name the column `state`.

5. Create a new column `ratio` whose value is the ratio `dist`/`speed`. Then compute the average and standard deviation of that column.

```{r Question 3}
data(cars)
print(head(cars,5))
set.seed(0)
state = sample(1:3, dim(cars[1]), replace = TRUE)
state = factor(state)
levels(state) = c('NY','CA','CT')
state
#print(head(cars2), 5)
ratio = cars['dist']/cars['speed']
cars2 = cbind(cars, state)
cars2 = cbind(cars2, ratio)
colnames(cars2) = c('dist','speed','state','ratio')

print(head(cars2))
print(mean(cars2['ratio']))


```


## Question #4: 

Read the `TimesSquareSignage.csv` and import it into R as `ts_data`. Then check the following features of the dataset:

1. The number of observations and the number of variables.

2. The type (`class`) of each variable.

3. How many missing values are there in the dataset?

4. Which rows (people) have missing values? Which columns (variables) include missing values?

```{r Question 4}
ts_data = read.csv('https://s3.amazonaws.com/graderdata/TimesSquareSignage.csv', stringsAsFactors=FALSE) #reading file in
print(head(ts_data)) 
cat(dim(ts_data)[1], 'rows  : ', dim(ts_data)[2], 'fields \n \n') #1. num obervations and variables
c = 1:18
#print(sapply(ts_data, class)) # 2. getting class of each field
column_na = sapply(c, na_count <- function(x) {# 3.getting na in each column
  sum(is.na(ts_data[,x]))
})
row_na = sapply(1:dim(ts_data)[1], na_count <- function(x){# 4. row_na, replaced list input
  sum(is.na(ts_data[x,]))
})

cat('na count in each vector =', column_na, '\n', '\n') #counting number of na values in each vector
#cat('na in each row =', row_na, '\n', '\n') # counting number of na values in each row
total_na = sum(column_na)
cat('total na in df =', total_na, '\n \n')

# adapting func to give bool if there are na over cols
is_na_col = sapply(1:dim(ts_data)[2], has_na <- function(x){
       any(is.na(ts_data[,x])) 
}
)
# generating bool vector for rows 
is_na_row = sapply(1:dim(ts_data)[1], has_na <- function(x){
       any(is.na(ts_data[x,])) 
}
)
# boolean filtering to give cols with na

cat('following colums contain missing data: \n', colnames(ts_data[is_na_col])) #returning col names of vectors with missing data
print(head(ts_data[is_na_row,])) #returns rows with na data

```

## Question #5:

From the Time Square dataset, we'd like to extract specific information about advertising in Midtown Manhattan. Obtain the following data frames and save them in a subfolder named `data` in your current directory as CSV files:

1. Observations from Upper Broadway. Save as `UpperBway.csv`.

2. Observations with greater-than-average square footage. Save as `SF.csv`.

3. The name, address, and location of the ten signs with the largest total square footage. Save as `TopTen.csv`. 

```{r Question 5}

#dir.creat('data') 
B_way = subset(ts_data, grepl("Broadway", ts_data$'Building.Address')) #grepl to obtain addresses w/ bway
print(head(B_way))
write.csv(B_way, 'data/UpperBway.csv') #run only once for HW
avg_SF = mean(as.numeric(ts_data[,'TOTAL.SF']), na.rm = TRUE) #only with avail. data
print(avg_SF)
SF = ts_data[ts_data['TOTAL.SF'] > 2104.7,] #watch out for the LAST comma in slicing
# this selects rows which contain "total.sf" greater than value,COMMA, all columns
write.csv(SF, 'data/SF.csv')
sf_order = order(-ts_data['TOTAL.SF']) # I could probably do this in one line
ts_bysize = ts_data[sf_order,]
TopTen = ts_bysize[1:10,c(1,2,3)]
write.csv(TopTen, 'data/TopTen.csv')

```




